---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Simple is beautiful. Just load the data as `dataset`
```{r}
dataset <- read.csv("activity.csv")
```
Prepare a function which converts interval number,
which is consist of hour and minute digits to actual time.

(Note: You may not use interval number as interger
since the interval between interval number 1059 and 1100,
which represent 10:59 and 11:00, is recognized as 41.)
```{r}
interval2time <- function(x) {
    strptime(sprintf("%02d:%02d", x%/%100, x%%100), format = "%H:%M")
}
```

## What is mean total number of steps taken per day?
Sum the steps for each day and store the variable `dailySteps`.
```{r}
dailySteps <- tapply(dataset$steps, dataset$date, sum)
hist(dailySteps, ylab = "Days")
```

The mean and median of daily steps are as follows.
```{r}
mean(dailySteps, na.rm = TRUE)
median(dailySteps, na.rm = TRUE)
```

## What is the average daily activity pattern?

```{r}
stepsPerInterval <- aggregate(steps ~ interval, data=dataset, mean)
stepsPerInterval$timeOfDay <- interval2time(stepsPerInterval$interval)
with(stepsPerInterval, plot(timeOfDay, steps,type = "l"))
stepsPerInterval[which.max(stepsPerInterval$steps),]
```

## Imputing missing values
At first, Let's investigate the total number of missing values.
```{r}
table(factor(is.na(dataset$steps)))
```
There are 2304 missing values out of 17568 observations.

Before devising a strategy to fill out missing values,
Let's Look at the distribution of missing value(NA)s.
The daily distribution of NA's are as follows.
```{r}
with(dataset, tapply(is.na(steps), date, sum))
```

As you can see, some days have no data.
Since 288 NA's means all the 24 * 60 / 5 = 288 data are NA's,
we cannot fill the NA as daily mean or median.

Instead, we can fill the NA's with daily mean of each interval.
Let's reuse the `stepsPerInterval` variable.
```{r}
names(stepsPerInterval)[2] <- "default"
combo <- merge(dataset, stepsPerInterval)
combo$steps[is.na(combo$steps)] <- combo$default[is.na(combo$steps)]
combo$default <- NULL
```

After fill out the missing values,
let's re-calculate the distribution of daily steps.
```{r}
dailyStepsFilled <- tapply(combo$steps, combo$date, sum)
hist(dailyStepsFilled, ylab = "Days")
mean(dailyStepsFilled, na.rm = TRUE)
median(dailyStepsFilled, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?
First, label the dates with weekday or weekend.
```{r}
date2daytype <- function(x) {
    wd <- as.POSIXlt(x)$wday
    if (wd == 0 | wd == 6) {
        "weekend"
    } else {
        "weekday"
    }
}
dataset$daytype <- sapply(dataset$date, date2daytype)
```

Take the average for each time interval by weekdays or weekends.
```{r}
stepsTrend <- aggregate(steps ~ daytype + interval,
                        data = dataset, mean, na.rm=TRUE)
stepsTrend$timeOfDay <- interval2time(stepsTrend$interval)
```

Draw the graph of the steps trend.
```{r}
library(ggplot2)
qplot(timeOfDay, steps, data=stepsTrend, facets=daytype ~ ., geom='path')
```

As you can see, there are some differences between weekdays and weekends.